The Nature of Breath: A Biomimetic Kinetic Sculpture that Breathes

### 2021-12-16 finalproject/report/ms1report.tex
small update to bib
close to done

Perhaps the most natural indicator of life is the movement of breathing. Respiration is an indicator of life at all scales - from single celled organisms to human beings. Whether involving oxygen in the conventional sense of breath, or operating at the cellular level, the process of energy production -- often for the purpose of movement -- is essential to life as we know it. Taking breath, the most natural of human movements, into close consideration, this work examines the relationship between what is natural and what is, or has been, alive. This mechanism may breathe in a natural way, but its appearance clearly identifies it as something which is purely artificial. There is no question that this breathing work does not actually posses life.

@@

While not the most complex respiratory system on the planet, the human respiratory system is very complicated.\cite{gasexchangers} To simplify this system for translation to a constructable mechanical structure, I broke it down into three main sub-systems. Each sub-system exists solely for the purpose of generating breath noises and not for the actual exchange of oxygen as would normally occur during respiration.

@@

\begin{figure}[h]
\includegraphics[width=0.5\textwidth]{images/buckettest.png}
\caption{Passesby observing a breathing bucket.}
\end{figure}

@@

The complexity of the human respiratory system lies not in its gross mechanics, but in the materials and molecular mechanics that allow for the transfer of oxygen to the blood. The mechanical goal of this work was to isolate the components of this complex system which form the noise created by breathing. These investigations revealed that breath is a sensitive topic and can be interpreted as an indicator of many things. In addition my intent to use breath as an indicator of life, I found that it is also a subtle and powerful indicator of emotion and health. The sound of breath can be used to great effect, however, these investigations highlighted that the movement of breath is a much more intuitive signifier of life.

@@

\bibliographystyle{plain}
\bibliography{ms1report}

>>>>

Perhaps the most natural indicator of life is the movement of breathing. Respiration is an indicator of life at all scales - from single celled organisms to human beings. Whether involving oxygen in the conventional sense of breath, or operating at the cellular level, the process of energy production -- often for the purpose of movement -- is essential to life as we know it.\cite{zimmer} Taking breath, the most natural of human movements, into close consideration, this work examines the relationship between what is natural and what is, or has been, alive. This mechanism may breathe in a natural way, but its appearance clearly identifies it as something which is purely artificial. There is no question that this breathing work does not actually posses life.

@@

The complexity of the human respiratory system lies not in its gross mechanics, but in the biological materials and molecular mechanics that allow for the transfer of oxygen to the blood.\cite{gasexchangers} To simplify this system for translation to a constructable mechanical structure, I broke it down into three main sub-systems. Each sub-system exists solely for the purpose of generating breath noises and not for the actual exchange of oxygen as would normally occur during respiration.

@@

\begin{figure}[h]
\includegraphics[width=0.5\textwidth]{images/buckettest.png}
\caption{Passerby observing a breathing bucket.}
\end{figure}

@@

The primary objective of this work was to create a kinetic sculpture which appeared to be alive and breathing, thereby asking what it means for an non-living, artifical construct to breathe in a natural, life-like way. This objective was approached trough a combination of auditory and mechanical explorations. These investigations revealed that breath is a sensitive topic and can be interpreted as an indicator of many things. In addition to my intent of using breath as an indicator of life, I found that it is also a subtle and powerful indicator of emotion and health. The ongoing COVID-19 pandemic has certainly made us more attuned to the subtle changes in breath that signal respiratory distress.

The sound of breath can be used to great effect, however, these investigations revealed that the movement of breath is a much more intuitive signifier of life. The mechanical goal of this work was to isolate the components of the respiratory system which form the noise created by breathing and highlight their movement. The intensity of viewer's reactions to the mechanical system developed confirmed that the movement is just as powerful as sound in indicating life.

@@

\bibliographystyle{unsrt}
\bibliography{ms1report}
